---
title: "Localize Campaign Analysis"
output: html_notebook
---

This is my analysis of Localize Marketing Campaigns, per the overview provided in the pdf instructions

### Read in the data
```{r}

library(tidyverse)
library(googlesheets4)


localize_user_sessions <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1kngeraxAiD_0j6MOO5vgVoNoUnU5UVEwfAyJBhDnFTk/edit#gid=0",range="user_sessions")

localize_user_sources<-googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1kngeraxAiD_0j6MOO5vgVoNoUnU5UVEwfAyJBhDnFTk/edit#gid=0",range="user_sources")

localize_campaign_costs<-googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1kngeraxAiD_0j6MOO5vgVoNoUnU5UVEwfAyJBhDnFTk/edit#gid=0",range="campaign_costs")


skimr::skim(localize_user_sessions)

```

```{r}

user_summary<-localize_user_sessions %>% 
  group_by(user_id) %>% 
  summarise(sessions=n(),
            days_visited=n_distinct(date),
            first_visit=min(date),
            latest_visit=max(date),
            time_between_first_last=difftime(latest_visit,first_visit,units="days")) 

user_summary%>% 
  ggplot(.,aes(sessions)) +
  geom_histogram(col="#003A42",
                 fill="#faf5f0",
                 binwidth = 1) +
  labs(title="Sessions per User",
        x ="Sessions", y = "Users")

```


```{r}

user_summary%>% 
  filter(sessions>1) %>% 
  ggplot(.,aes(time_between_first_last)) +
  geom_histogram(col="#003A42",
                 fill="#faf5f0",
                 binwidth = 1) +
  labs(title="Days Between First and Last Session",
        x ="Days", y = "Users")



```

Let's take a look at our secondary datasource,

```{r}

skimr::skim(localize_user_sources)


localize_user_sources %>% 
  group_by(campaign_source) %>% 
  summarise(users=n_distinct(user_id),
            occurences=n())

```

```{r}

skimr::skim(localize_campaign_costs)

```



A cursory overview of the campaign costs and user_sources data shows a mismatch of a few campaign names.  _For the puroses of this exercise, I wil assume the instagram and google campaigns are intended to be referencing the same campaigns._  I'm going to add a new column to our campaign costs dataset to account for the name differences

```{r}

localize_campaign_costs_edited <- localize_campaign_costs %>% 
  mutate(campaign_edited=case_when (
    campaign=='google_adwords'~'google_1',
    campaign=='instagram_thx'~'insta_1',
    TRUE~campaign
  ),
  ad_platform=case_when(
    grepl('google',campaign,ignore.case = TRUE)~'Google',
    grepl('instagram',campaign,ignore.case = TRUE)~'Instagram',
    grepl('fb_',campaign,ignore.case =TRUE)~'Facebook',
    TRUE~'unknown'
  ))


```




```{r}

#let's join the campaign source and cost data to the dataset

localize_session_full<-localize_user_sessions %>% 
  left_join(localize_user_sources) %>% 
  left_join(localize_campaign_costs_edited,by=c("campaign_source"="campaign_edited")) %>% 
  select(-is_session,-campaign_source) 

localize_session_full %>% 
  group_by(campaign) %>% 
  summarise(users=n_distinct(user_id),
            sessions=n(),
            spend=sum(ppu),
            spu=sessions/users)

```

Let's assume the cost is associated at the User, not click level - what does that look like for a cost per session then

```{r}


#this is sort of what we want, but don't necessarily want to lose the campaign name for each row
localize_session_full<-localize_user_sessions %>% 
  arrange(date) %>% 
  group_by(user_id) %>% 
  mutate(join_key=row_number()) %>% 
  left_join(localize_user_sources) %>% 
  left_join(localize_campaign_costs_edited %>% 
              mutate(join_key=1),by=c("campaign_source"="campaign_edited","join_key"="join_key")) %>% 
  select(-is_session,-campaign_source) 


#this is a really lazy way to do this but will work for our purposes
localize_session_full<-localize_user_sessions %>% 
  arrange(user_id,date) %>% 
  group_by(user_id) %>% 
  mutate(sess_num=row_number()) %>% 
  left_join(localize_user_sources) %>% 
  left_join(localize_campaign_costs_edited,by=c("campaign_source"="campaign_edited")) %>% 
  mutate(ppu=ifelse(sess_num==1,ppu,0)) %>% 
  select(-is_session,-campaign_source) 


localize_session_full %>% 
  group_by(campaign) %>% 
  summarise(users=n_distinct(user_id),
            sessions=n(),
            spend=sum(ppu),
            spu=sessions/users,
            cost_per_sess=spend/sessions)



```

```{r}

localize_session_full %>% 
  group_by(date, campaign) %>% 
  summarise(sessions=n()) %>% 
  ggplot(aes(x=date, y=sessions, group=campaign,colour=campaign)) +
  geom_line()+
  labs(title="Weekly Accumulating Leads by Lifecycle Stage", x="Week Starting", y="Count of Leads") +
    theme(legend.title=element_blank(),legend.text=element_text(size=5),legend.position='bottom') 


```

